Using Scipion for Stream Image Processing at EM Facilities

### 2018-03-22 2.overall.tex
before class

\scipion is a flexible framework that allows to create different workflows by choosing among many algorithms at each step. In this work we do not recommend a particular workflow  over others since we realize that different specimens, microscopes or facilities may have different requirements. Nevertheless, since a use case may clarify \scipion capabilities, in the following we describe the pipeline offered by default at the CNB cryoEM Service.

>>>>

\scipion is a flexible framework that allows to create different workflows by choosing among many algorithms at each step. In this work we do not recommend a particular workflow  over others since we realize that different specimens, microscopes or facilities may have different requirements. Nevertheless, since a use case may clarify \scipion capabilities, in the following we describe the pipeline offered by default at the \emph{Centro Nacional de Biotecnología} (CNB) cryoEM Service.

### 2018-03-22 3.workflow.tex
before class

 \item \emph{Ume\aa\ Core Facility for Electron Microscopy} (UCEM, \url{http://www.kbc.umu.se/english/ucem})[** Jose Miguel can you verify this?] 
 \item \emph{Science for life Laboratory} (SciLifeLab, \url{http://www.scilifelab.se})
 \item \emph{ESRF Cryo-Electron Microscope} (\url{http://www.esrf.eu/home/UsersAndScience/Experiments/MX/About_our_beamlines/CM01.html})
 \item \emph{Electron Bio-Imaging Centre at Diamond Light Source} (eBIC, \url{http://www.diamond.ac.uk/Science/Integrated-facilities/eBIC.html})
 \item \emph{CNB CryoEM Service} (\url{http://www.cnb.csic.es/index.php/en/research/core-facilities/microscopia-crioelectronica})
 \item \emph{Molecular Microscopy Consortium} (\url{https://www.niehs.nih.gov/research/atniehs/facilities/mmc/index.cfm})[**Pablo can you confirm this?]
 \item \emph{National Cancer Institute - NIH} (\url{https://cmm.nci.nih.gov/})[**Pablo can you confirm this?]

\end{itemize}

\scipion original design satisfied the requirements of the CNB, in which \scipion automatically fetches newly recorded movies from a network mounted disk. After this first installation, \scipion was deployed on SciLifeLab (a centralized facility located at Stockholm) where a rigorous booking system is followed. There, \scipion was adapted to interface with this booking platform in order to produce better reports. Lately,  \scipion has been installed in  large synchrotrons  such as ESRF and Diamond  where projects are handled by ISPyB (a customized laboratory management information system -LIMBS- \citep{Delageniere2011}) and data are saved on a distributed file system that heavily penalizes access to lists of files. In this environment, \scipion needs to constantly interchange information with ISPyB, and it cannot longer check if there is a new file in a particular storage device but must wait to be informed by the system that a new file has been created. All these particularities, that will be described in detail in the following, could only be handled thanks to a careful design that prioritizes flexibility at several levels.

@@

\subsection{What will do in future}
%Particle picking
%Extract particle
%2D classify (either classes or identify outliers)
%Initial model
%More complex reports
[** this section may disappear]

>>>>

 %\item \emph{Ume\aa\ Core Facility for Electron Microscopy} (UCEM, \url{http://www.kbc.umu.se/english/ucem})[** Jose Miguel can you verify this?] 
 \item \emph{Science for life Laboratory} (SciLifeLab, \url{http://www.scilifelab.se}). Both at Stockholm and Ume\aa\ facilities)
 \item \emph{ESRF Cryo-Electron Microscope} (\url{http://www.esrf.eu/home/UsersAndScience/Experiments/MX/About_our_beamlines/CM01.html})
 \item \emph{eBIC at Diamond Light Source} ( \url{http://www.diamond.ac.uk/Science/Integrated-facilities/eBIC.html})
 \item \emph{CNB CryoEM Service} (\url{http://www.cnb.csic.es/index.php/en/research/core-facilities/microscopia-crioelectronica})
 \item \emph{Molecular Microscopy Consortium - NIH} (\url{https://www.niehs.nih.gov/research/atniehs/facilities/mmc/index.cfm})[**Pablo can you confirm this?]
 \item \emph{National Cancer Institute - NIH} (\url{https://cmm.nci.nih.gov/})[**Pablo can you confirm this?]

\end{itemize}
and it is being considered in many more.

\scipion original design satisfied the requirements of the CNB, in which \scipion automatically fetches newly recorded movies from a network mounted disk. After this first installation, \scipion was deployed on SciLifeLab (a centralized facility located at Stockholm) where a rigorous booking system is followed. There, \scipion was adapted to interface with this booking platform in order to produce better reports. Lately,  \scipion has been installed in  large synchrotrons  such as ESRF and Diamond  where projects are handled by ISPyB (a customized laboratory management information system -LIMBS- \citep{Delageniere2011}) and data are saved on a distributed file system that heavily penalizes access to lists of files. In this environment, \scipion needs to constantly interchange information with ISPyB, and it cannot longer check if there is a new file in a particular storage device but must wait to be informed by the system that a new file has been created.

In the following, through three use cases, we describe in detail how \scipion has been adapted to the different environments.

%All these particularities, that will be described in detail in the following, could only be handled thanks to a careful design that prioritizes flexibility at several levels.

@@

\subsection{\scipion near future}
%Particle picking
%Extract particle
%2D classify (either classes or identify outliers)
%Initial model
%More complex reports
plug-in style, independend

### 2018-03-22 3.workflow_cnb.tex
before class

\subsection{CNB}

This section in divided in two parts. In the first one we discuss the network setup that allow us to move the data from the microscope to the user laboratory. Then  we comment on an additional software (called  \emadmin) that connects \scipion with the microscope acquisition 
software (EPU) and records the facility activity.

\subsubsection{Network Setup}

@@

Data sets are recorded using EPU. The microscope control system has a 70 TB NAS (Network-attached storage) unit for saving movies as they are obtained (we will refer to this NAS as microscope-NAS). This storage unit is shared through a Samba share and connected to the private network using a 10Gb connexion. \scipion is executed on a Linux server which has two 1Gb NICs (network interface controllers) that connect it to the private and the LAN networks respectively (we will refer to this machine as \scipionbox). \scipionbox has 32 CPUs at 2.40GHz, 62 Gb of RAM memory and 2 \textit{Quadro M4000} GPUs. Finally, a second NAS with 4 1Gb NICs is linked to the private, LAN and WAN networks (we will refer to this second NAS as output-NAS).

The data flow is a follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox will temporary read the movies from 
the microscope-NAS but will not store them locally, (c) each hour, output-NAS will connect to \scipionbox and copy or update the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup the \scipion project and the movies to a USB disk or serve this information through the Internet and, (e) in parallel, \scipionbox transfers to  output-NAS different html reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox}).

Although we try to discourage the use of external USB hard disks versus network as way of retrieving the raw and processing data, this is the preferred option for more of our users.
USB 3.0 drives can theoretically handle data flows of 300 MB/s, but our experience is that, in this busy environment, many of them  have trouble exceeding 50 MB/s. In most cases, 50 MB/s is enough to transfer the data before the following microscope shift starts. Nevertheless, real-world is tough and it happens that two, or even three,  users try to download their data simultaneously. Fortunately, output-NAS works as front-end and isolates the data acquiring and processing infrastructure from the user downloading. Network users access to output-NAS using a chrooted common account. This account only allows rsync connections and make impossible to retrieve data if the name of the project to be downloaded is unknown. In order to speed up transfer weak ciphers (as \emph{arcfour}) has been activated in output-NAS.

@@

\emadmin was born as a script that created a normalized tree of directories  in which to store the movies acqured by the microscope and help users to choose between a few predetermined image processing workflows. At present it was grown in a client-server application that in addition to the old script capacities it has incorporate the following functions:

@@

\emadmin is free software available at github repository (\url{https://github.com/rmarabini/webservices/tree/master/EMadmin}). You are welcomed to try it but the software is provided as it is and no support or maintenance service is offered.

>>>>

\subsection{Case 1: Scipion at CNB}
The National Center for Biotechnology (CNB) forms part of the Spanish National Research Council (CSIC), the largest public research institution in Spain. 
The CNB Cryo-Electron Microscopy Facility is a joint effort of the Centro Nacional de Biotecnolog\'{i}a (CNB-CSIC) and the Centro de Investigaciones Biol\'{o}gicas (CIB-CSIC). The mission of the facility is to provide researchers access to the latest cryo-EM technology for high resolution imaging.

This section in divided in two parts. In the first one we discuss the network setup that allow us to move the data from the microscope to the user laboratory. Then  we comment on an additional software (called  \emadmin) that connects \scipion with the microscope acquisition software (\epu) and records the facility activity.

\subsubsection{Network Setup and IT Infrastructure}

@@

Data sets are recorded using \epu. The microscope control system has a 70 TB NAS (Network-attached storage) unit for saving movies as they are obtained (we will refer to this NAS as microscope-NAS). This storage unit is shared through a Samba share and connected to the private network using a 10Gb connexion. \scipion is executed on a Linux server which has two 1Gb NICs (network interface controllers) that connect it to the private and the LAN networks respectively (we will refer to this machine as \scipionbox). \scipionbox has 32 CPUs at 2.40GHz, 62 Gb of RAM memory and 2 \textit{Quadro M4000} GPUs. Finally, a second NAS with 4 1Gb NICs is linked to the private, LAN and WAN networks (we will refer to this second NAS as output-NAS).

The data flow is a follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox will temporary read the movies from 
the microscope-NAS but will not store them locally, (c) each hour, output-NAS will connect to \scipionbox and copy -or update- the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup \scipion project and the movies to a USB disk or serve this information through the Internet and, (e) in parallel, \scipionbox transfers to  output-NAS different html reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox}).

Although CNB staff try to discourage the use of external USB hard disks versus network as way of retrieving the raw and processing data, this is the preferred option for most users.
USB 3.0 drives can theoretically handle data flows of 300 MB/s, but CNB experience is that, in this busy environment, many of them  have trouble exceeding 50 MB/s. In most cases, 50 MB/s is enough to transfer the data before the following microscope shift starts. Nevertheless, real-world is tough and it happens that two, or even three,  users try to download their data simultaneously. Fortunately, output-NAS works as front-end and isolates the data acquiring and processing infrastructure from the user downloading. Network users access to output-NAS using a chrooted common account. This account only allows rsync connections and make impossible to retrieve data if the name of the project to be downloaded is unknown. In order to speed up transfer weak ciphers (as \emph{arcfour}) has been activated in output-NAS.

@@

\emadmin was born as a script that created a normalized tree of directories  in which to store the movies acquired by the microscope, helped users to choose between a few predetermined image processing workflows and finally launched \scipion. At present it was grown in a client-server application that in addition to the old script capacities it has incorporate the following functions:

@@

\emadmin is free software available at github repository (\url{https://github.com/rmarabini/webservices/tree/master/EMadmin}). You are welcomed to try it but be aware that \emadmin is designed to be used at CNB and therefore, no support is provided for it.

### 2018-03-22 3.workflow_scilifelab.tex
before class

\subsection{Swedish National CryoEM Facility}

The Swedish National Cryo-EM Facility offers access to state-of-the-art equipment and expertise 
in single particle cryo-EM and cryo electron tomography (cryo-ET). The Facility has two nodes: at SciLifeLab 
in Stockholm and at Umeå University. SciLifeLab in Stockholm offers single-particle service with a 
Talos Arctica for sample optimisation and a Titan Krios for high-resolution data collection. 
The Umeå node is expected to become operational during 2018, and will offer cryo-ET with a 
Titan Krios and a Scios DualBeam SEM.

The facilities can be accessed by Swedish researchers through a peer-reviewed routine. Applications can be 
submitted through an Application Portal [**reference???] and will be evaluated once every three months based on their 
scientific merit and technical feasibility by a national Project Evaluation Committee. On the other side, 
some time of the facility is reserved for internal research groups at Stockholm University. Internally, the allocation
of the microscope (and other instruments) time is managed by a Booking System [** name of the booking system and reference???].

At the beginning of 2016, the Cryo-EM facility at SciLifeLab became one of the early adopters of \scipion 
for on the fly data processing. Before, there were some home-made scripts to perform motion correction (with motioncor)
and CTF estimation (with Gctf). The microscope operators recognized the importance of  
a more general framework with the possibility to be extended with new programs. Another strong point was the 
the availability of graphical tools for data analysis and quick feedback about the data collection. Initially, the 
same setup script from the CNB was used, but it had some limitations for this more complex setup with more 
requirements. In the following sections we will briefly describe the computational infrastructure at the
SciLifeLab node and the implementation of a Session Wizard to automate the initial setup while fetching 
information from other sources external to Scipion.

\subsubsection{Infrastructure}
At the heart of the SciLifeLab computational infrastructure is a storage and pre-processing server (the
staging server). This machine has roughly 200 TB of storage (4 ZFS RAIDZ2 pools
with 11 HDDs each), 2 NVIDIA GeForce GTX 1070 GPUs, 2 Intel Xeon E5-2630v4 CPUs
(10 cores, 2.2 GHz, HT), 384 GB RAM, and a dual-port 10 GbE network card. The
storage pool is exported via NFS and Samba, so the microscope computers can
write data there directly and other machines can access it.

Unfortunately the K2 computer writes data to a local SSD RAID, the Falcon-III to
a dedicated storage server (the offload server), in order to avoid interruption
of data collection due to network issues. So when users start data collection,
they start a script that continously moves the produced data to the staging
server.

By having all data on the staging server, in an organized folder structure, we
can run pre-processing directly on the server. If processing is started soon
after data collection the incoming file data should still be in RAM and does not
have to be read from disk, which is important with performance in mind. Users
can copy their data to external USB drives from two workstations in the
microscope room, that have access to the storage pool via NFS.
There is also a dedicated download server with access to the NFS export,
reachable from the Internet. Here users can log in to a restricted shell that
allows downloads via rsync (over SSH) or SFTP.

All these computers are conntected to a 10 GbE network switch. This is a private
network without external access. The only machine accessible from the Internet
is the data download server, which has a dual-port 10 GbE network adapter (one
used for the private network, one for the external network). Both the K2
computer and the Falcon-III offload server had one unused 10GbE interface that
could be used to connect them to our private network, without having to
interfere with the default setup.

\subsubsection{Session Wizard}

The initial attempt to ease the setup of \scipion projects for stream processing was based  
on the original script developed at the CNB. But it was difficult to reuse the existing code due to a different folder
structure and, more important, a different logic about users and microscopes management. To address the new set
of requirements, a new ``wizard'' is being developed at SciLifeLab and can be found in the following open source repository: 
\url{https://github.com/delarosatrevin/scipion-session}

For this new wizard, a basic data model was designed representing the current processing and data 
organization at SciLifeLab. The model is composed by the following entities:

@@

When an user starts a data collection, the wizard is also executed using the command \textit{session-titan} 
(or \textit{session-talos}). The wizard will fetch the reservations for this day using a webservices API provided by 
the Booking System and will check wich user is operating the microscope. From the reservation information, 
the wizard will also detect if it is an internal project or a national facility project. If the latter,
more information will be retrieved from the Applicacion Portal (such as principal investigator, project code, 
contact person or invoice address). After gathering all this information, the wizard setup the Scipion project
for streaming processing. For example, the movies-import protocol will already point to the location where
the files will be written for this session. Moreover, taking into account which microscope-camera is being 
used, the files pattern can be inferred.

The current implementation of the wizard stores each session in a simple database file. This database, together
with extra information from the Booking System and the Appplication Portal, is used to generate reports for 
a given period with the associated invoices. The generation of a session report is under development, that 
will provide a summary and useful information for users to access the acquired data. The existing data model 
represents the specify usage at the SciLifeLab facility and there are not inmediate plans to make it more 
general and easy to reuse. Nonetheless, the model is quite simple and should to be difficult to modify it to fit 
other requirements.

>>>>

\subsection{Case 2: Scipion at SciLifeLab}

The Swedish National Cryo-EM Facility offers access to state-of-the-art equipment and expertise in single particle cryo-EM and cryo electron tomography (cryo-ET). The Facility has two nodes: at SciLifeLab in Stockholm and at Ume\aa\ University. SciLifeLab in Stockholm offers single-particle service with a 
Talos Arctica for sample optimisation and a Titan Krios for high-resolution data collection. The Ume\aa\ node is expected to become operational during 2018, and will offer cryo-ET with a Titan Krios and a Scios DualBeam SEM.

The facilities can be accessed by Swedish researchers through 
a peer-reviewed process. Applications can be submitted through an Application Portal (\url{https://cryoem.scilifelab.se/}). % and will be evaluated once every three months based on their scientific merit and technical feasibility by a national Project Evaluation Committee. %On the other side, some time of the facility is reserved for internal research groups at Stockholm University. 
Internally, the allocation of the microscope (and other instruments) time is managed by a Booking System [**ROB-JM name of the booking system and reference???].

At the beginning of 2016, the Cryo-EM facility at SciLifeLab became one of the early adopters of \scipion for on the fly data processing. Before, there were some home-made scripts to perform motion correction (with motioncor) and CTF estimation (with Gctf). The microscope operators recognized the importance of  
a more general framework with the possibility to be extended with new programs. Another strong point was the the availability of graphical tools for data analysis and quick feedback about the data collection. Initially, the same setup script used at CNB was adopted, but it did not fully satisfy the facility requirements and a new one has been developed. In the following sections we will briefly describe the computational infrastructure at the SciLifeLab node [** The one in  Stockholm?] and the implementation of a Session Wizard to automate the initial setup while fetching information from other sources external to \scipion.

\subsubsection{Network Setup and IT Infrastructure}
At the heart of the SciLifeLab computational infrastructure is a storage and pre-processing server (the
staging server). This machine has roughly 200 TB of storage (4 ZFS RAIDZ2 pools
with 11 HDDs each), 2 NVIDIA GeForce GTX 1070 GPUs, 2 Intel Xeon E5-2630v4 CPUs
(10 cores, 2.2 GHz, HT), 384 GB RAM, and a dual-port 10 GbE network card. The
storage pool is exported via NFS and Samba, so the microscope computers can
write data there directly and other machines can access it.

Unfortunately, the K2 [** Jose Miguel, what is K2?, the computer described in the previous paragraph?] computer writes data to a local SSD RAID while the Falcon-III does it to a dedicated storage server (the offload server). %, in order to avoid interruption of data collection due to network issues. 
So when users start the data collection, they launch a script that continuously moves the produced data to the staging server.

[** ROB. I do not understand this sentence, what preprocessing:  By having all data on the staging server, in an organized folder structure, we
can run pre-processing directly on the server.] If data processing is started soon
after data collection the incoming data file should be in RAM-cache and  no
disk access in needed, which is important with performance in mind. Users
can copy their data to external USB drives from two workstations in the
microscope room, that have access to the storage pool via NFS.
There is also a dedicated download server with access to the NFS export,
reachable though the Internet. Here users can log in to a restricted shell that
allows downloads via rsync (over SSH) or SFTP.

All these computers are connected to a 10 GbE network switch. This is a private
network without external access. The only machine accessible from the Internet
is the data download server, which has a dual-port 10 GbE network adapter (one
used for the private network, one for the external network). Both the K2
computer and the Falcon-III offload server had one unused 10GbE interface that
could be used to connect them to our private network, without having to
interfere with the default setup.

\subsubsection{Session Wizard}
In order to interface \scipion with the microscope infrastructure an upgraded version of the original script developed at the CNB was used. Unfortunately, it was difficult to reuse the existing code due to a different folder structure and, more important, a different logic about users and microscopes management. To address the new set of requirements, a new wizard is under development at SciLifeLab. For this new wizard, a basic data model was designed representing the current processing and data organization at SciLifeLab. The model is composed by the following entities:

@@

When an user starts a data collection, the wizard is also executed using the command \textit{session-titan} (or \textit{session-talos}). The wizard will fetch the reservations for this day using a webservices API provided by the Booking System and will check which user is operating the microscope. From the reservation information, the wizard will also detect if it is an internal project or a national facility project. If the latter, more information will be retrieved from the Application Portal (such as principal investigator, project code, contact person or invoice address). After gathering all this information, the wizard setup the \scipion project for streaming processing. %For example, the movies-import protocol will already point to the location where the files will be written for this session. Moreover, taking into account which microscope-camera is being  used, the files pattern can be inferred.

The current implementation of the wizard stores each session in a simple database file. This database, together with extra information from the Booking System and the Appplication Portal, is used to generate reports for a given period with the associated invoices. The generation of a session report is under development, that 
will provide a summary and useful information for users to access the acquired data.

The described session wizard can be found at  \url{https://github.com/delarosatrevin/scipion-session}. It has been design to satisfy SciLifeLab requirements and, therefore, although readers are welcomed to try it, no support is offered.

%existing data model represents the specify usage at the SciLifeLab facility and there are not immediate plans to make it more general and easy to reuse. Nonetheless, the model is quite simple and should to be difficult to modify it to fit other requirements.

### 2018-03-22 5.conclusions.tex
before class

\section{Discussion}
[**TODO]: Discuss the obtained numbers in the result section
[**REmove this section?]

@@

 In a workflow approach, users (including microscope staff) can reasonably be expected to edit the paths. They can potentially play with the pipeline in ways that aren't possible when having a programmer write code for them. Furthermore, not having the processing path hardcoded in the program means you don't have to recompile, so changes can be quickly tested   even at run time. The \scipion workflow approach provides provenance, that is,  for any created object (micrograph, CTFs, class, etc) \scipion stores metadata such as which programs, which parameters, and which exact path was followed to create it.  In this way,  real tracking and reproducibility are provided.

@@

\scipion is provided 
freely as open source software. Online documentation
describing \scipion download and installation is available
at \url{http://scipion.cnb.csic.es/m/download_form/} and \url{https://github.com/I2PC/scipion/wiki/How-to-Install}, respectively. Cloud installation instruction are available at \url{https://github.com/I2PC/scipion/wiki/Scipion-in-the-Cloud}. Other public software mentioned in this article as \emadmin, XXXX and YYYY are available at \url{https://github.com/rmarabini/webservices/tree/master/EMadmin}, www eee respectively.
If you would like more information about how to install \scipion or you have suggestions on how to improve it contact us at scipion@cnb.csic.es.

>>>>

%\section{Discussion}
%[**TODO]: Discuss the obtained numbers in the result section
%[**REmove this section?]

@@

 In a workflow approach, users (including microscope staff) can reasonably be expected to edit the paths. They can potentially play with the pipeline in ways that aren't possible when having a programmer write code for them. Furthermore, not having the processing path hardcoded in the program means you don't have to recompile, so changes can be quickly tested   even at run time. The \scipion workflow approach provides provenance, that is,  for any created object (micrograph, CTFs, class, etc) \scipion stores metadata such as which programs, which parameters, and which exact path was followed to create any output data.  In this way,  real tracking and reproducibility are provided.

@@

\scipion is provided 
freely as open source software. Online documentation
describing \scipion download and installation is available
at \url{http://scipion.cnb.csic.es/m/download_form/} and \url{https://github.com/I2PC/scipion/wiki/How-to-Install}, respectively. Cloud installation instruction are available at \url{https://github.com/I2PC/scipion/wiki/Scipion-in-the-Cloud}. Other public software mentioned in this article as \emadmin, \emadmin and \emph{wizard session} are available at \url{https://github.com/rmarabini/webservices/tree/master/EMadmin} and \url{\url{https://github.com/delarosatrevin/scipion-session}} respectively.
If you would like more information about how to install \scipion or you have suggestions on how to improve it contact us at scipion@cnb.csic.es.
